Functional Analysis of CGM Data

### 2022-12-02 StatNet/cgm_fda.qmd
Update cgm_fda.qmd

---
title: "Functional Analysis of CGM Data"
author: "Tim Vigers"
format: 
  revealjs:
    scrollable: true
editor: visual
---

@@

sleep = read.csv("C:/Users/timbv/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Raw/sleep.csv",

@@

# cleandata("C:/Users/timbv/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Raw/CGM",
#           "C:/Users/timbv/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",

@@

cgms = list.files("C:/Users/timbv/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",

@@

save(ydf,file = "C:/Users/timbv/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/R_Objects/ydf.Rdata")

@@

## Our Dataset {.smaller}

@@

```{r}
#| include: false
# load("C:/Users/timbv/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/R_Objects/fpca_sc_fit.Rdata")
```

@@

-   What is the best way to pick the order of basis functions, etc.? Visually?
-   Currently I'm forcing the glucose measures to be on the interval \[0,1\], which means that the period for the Fourier basis is "nighttime." Since those with shorter sleep periods will have fewer observations, I think this means that

>>>>

---
title: "Functional Analysis of CGM Data"
author: "Tim Vigers"
format: 
  revealjs:
    scrollable: true
    smaller: true
editor: visual
---

@@

sleep = read.csv("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Raw/sleep.csv",

@@

# cleandata("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Raw/CGM",
#           "~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",

@@

cgms = list.files("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/Data_Clean/CGM",

@@

save(ydf,file = "~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/R_Objects/ydf.Rdata")

@@

## What Is Functional Data Analysis (FDA)?

- Most statistics use some sort of parametric model $p(x|\boldsymbol{\theta})$.
  - We try to estimate the parameters of *p*.
- Assume the data are normally distributed $N(\mu,\sigma^2)$.
  - Can use the sample mean $\bar{x}$ as an estimate of the location parameter $\mu$.
- Sometimes we do not want to assume a density function in advance.
  - Interested in estimating $p$ itself.
    - What is the smooth function that generates our data?

## Goals and Uses

- The goals of FDA are essentially the same as other areas of statistics:
  1. Represent data to aid in analysis and to highlight characteristics
  2. Study patterns and sources of variation
  3. Explain variation in an outcome using independent variables

- Common uses of FDA:
  - Growth data
  - Imaging data 
  - Wearable device data
  - Accelerometry (physical activity) 
  - Heart rate
  - **Continuous glucose monitoring** (CGM)

## Semi- and Non-Parametric Regression

- The basic linear model is: $y_i= \beta_0+\beta_1 x_i+\epsilon_i$ for $1\leq i \leq n$
- Each predictor has a simple form
  - The overall shape is determined by the model rather than the data.
- The non-parametric version of this model is: $y_i = f(x_i)+\epsilon_i$

## Semi- and Non-Parametric Regression

- $f(x_i)$ is an arbitrary function.
  - The only constraint is that $f(x_i)$ must be a smooth function.
  - This model is much more flexible than parametric.
- A semi-parametric model essentially uses parametric models for some predictors, and non-parametric models for others.

## Basis Functions

- How do we represent $f(x)$?
- If $b_j(x)$ is the $j^{th}$ basis function, then: $$f(x)=\sum_{j=1}^k \beta_j b_j(x)$$
- Polynomials are a simple form of basis function: $$f(x)=\beta_1+\beta_2x+\beta_3x^2+\beta_4x^3+\beta_5x^4$$ and $$y_i=\beta_1+\beta_2x+\beta_3x^2+\beta_4x^3+\beta_5x^4+\epsilon_i$$

## Our Dataset

@@

```{r}
#| include: false
# load("~/Dropbox/Work/Vigers/CF/Christine Chan/CGM and FDA/R_Objects/fpca_sc_fit.Rdata")
```

@@

- I'm having some trouble running
